# Reproducible Research Assignment

Defining libraries
```{r echo=FALSE}
library(knitr)
library(markdown)
library(rmarkdown)
library(Hmisc)

```


##Loading and Preprocessing the data
1- Load the data
```{r}
activity <- read.csv(file = "activity.csv", header = TRUE, sep = ",")
names(activity)
str(activity)
```

##What is mean total number of steps taken per day?
1- Make a histogram of the total number of steps taken each day
```{r}

sumSteps <- aggregate(steps ~ date, data = activity,  FUN="sum")
sumSteps

hist(sumSteps$steps, col = 'blue',  xlab = 'Total Steps per Day', main = "Histogram of Activities")
```

2- Calculate and report the mean and median total number of steps taken per day
```{r}

meanStepsPerDay <- aggregate(steps ~ date, data = activity, FUN = "mean")
meanStepsPerDay
totalMeanSteps <- mean(sumSteps$steps)
totalMeanSteps

totalMedianSteps <- median(sumSteps$steps)
totalMedianSteps
```

##What is the average daily activity pattern?
1- Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
averageSteps <- aggregate(steps~interval, data=activity, FUN="mean")
averageSteps
```

Time series plot of the average number of steps taken
```{r}
plot(x = averageSteps$interval, y = averageSteps$steps, type = "l", lwd=2, xlab = "Intervals",
     ylab = "Average number of steps", main = "Average Daily activity Pattern")
```

2- Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of step
```{r}
averageSteps[which.max(averageSteps$steps),]
averageSteps[which.max(averageSteps$steps),]$interval
```
##Imputing Missing Values
1- Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
missingValues <- is.na(activity$steps)
table(missingValues)
```

2- Creating a New Dataset . Filling the missing data with average.
```{r}
newDataset <- activity
newDataset$steps <- with(newDataset,impute(steps,mean))
newDataset$date <- as.Date(newDataset$date)
str(newDataset)
sumNewdataset <- aggregate(steps ~ date, data = newDataset, FUN = "sum", na.rm = TRUE)
sumNewdataset
```

3- Histogram of the total number of steps taken each day after missing values are imputed.
```{r}
hist(sumNewdataset$steps, main="Steps Per Day By Frequency",xlab="Daily Steps")
```
4- Mean and Median of newDataset
```{r}
mean(sumNewdataset$steps, na.rm = TRUE)
median(sumNewdataset$steps, na.rm = TRUE)
```
##Are there differences in activity patterns between weekdays and weekends?
```{r}
dataWeek <- transform(newDataset, day = weekdays(date))
```
1-Create a new factor variable in the dataset with two levels ??? ???weekday??? and ???weekend??? indicating whether a given date is a weekday or weekend day.
```{r}
weekdays <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
weekdays
dataWeek$weekClass <- factor(dataWeek$day %in% weekdays, 
                          levels=c(FALSE, TRUE), labels = c("weekend", "weekday"))
```
Subsetting the Data in Weekdays and Weekend
```{r}
weekends <- dataWeek[dataWeek$weekClass=="weekend",]
weekendSum <- aggregate(steps ~ interval, data=weekends, FUN="sum")
weekdays <- dataWeek[dataWeek$weekClass=="weekday",]
weekdaySum <- aggregate(steps ~ interval, data=weekdays, FUN="sum")
```

2- Make a panel plot containing a time series plot (i.e.of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
par(mfrow=c(2,1))
plot(weekdaySum$interval, weekdaySum$steps, ylab = "Weekday Steps", xlab = "Weekdays Interval ", type="l", main="Weekday Activity")

plot(weekendSum$interval, weekendSum$steps, ylab = "Weekend Steps", xlab = "Weekends Interval", type="l", main="Weekend Activity")
```
